An Accurate Model for SARS-CoV-2 Pooled RT-PCR Test Errors

### 2021-09-12 paper.tex
almost done

\section*{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in the fight against the COVID-19 pandemic. PCR testing can be performed on individuals suspected of infection and hence break potential transmission chains. Moreover, in some countries, PCR testing has become a routine procedure of surveillance, testing all incoming air passengers in airports (ref), schools students (ref), or even the entire adult population (ref).  
As such, the need for large-scale testing has
resulted in the application of techniques to increase the throughput
of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
  MatrixPooling, LionDorfman, CherifReview}. These techniques, often
termed \emph{pooling}, or group testing, have originated in the 1940's
in the seminal work of Dorfman and his pooling technique
\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
selects $N$ individuals and performs a single RT-PCR test on their
combined (\emph{pooled}) samples. If the pooled test yields a positive
result, then each individual is retested separately; otherwise,
everyone is declared negative. The throughput efficiency of Dorfman
pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
its error rates thoroughly investigated \cite{Kim, Simplistic1,
  OptimalDorfmanPool}.

@@

\section*{Results}\label{section:results}
\subsection*{Scheme false-negative}
We plot $\err{exact}$ for varying prevalence $q$ and sensitivity $\Se$
values, and make the comparison with $\err{common}$. As recommended by
\cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
different prevalence values. We observe that for a false-positive rate
$\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
sensitivity and prevalence values \cite{KitComparison,
  InterpretingCOVID19Test, EstimatingRatesLourenco,
  FalsePositiveEstimate}, the increase in the scheme false-negative rate
(compared to $1-\Se$), expressed by $\err{exact}$, is at leaset
$60\%$ (Figure \ref{fig1}).

@@

\section*{Discussion}
In this study, we developed a more realistic and novel probabilistic
model for the outcomes of pooled RT-PCR tests, parametrized for
SARS-CoV-2. Contrary to the common assumption, we assume, based on
data (Table~\ref{table}), that multiple infected individuals increase
the likelihood of a positive pooled test result. A direct consequence
of our model is that low values of infection prevalence increase the
false-negative rates of Dorfman pooling. Importantly, our results
remain qualitatively similar under varying parameter values, in the
observed ranges \cite{KitComparison,EstimatingRatesKucrika,
  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
\ref{fig1}). Therefore, our results give rise to a conflict: low
infection prevalence leads to high efficiency of Dorfman pooling
\cite{DorfmanYuvalDor}, while also increasing false-negative rates.

>>>>

\section*{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in the fight against the COVID-19 pandemic. RT-PCR testing can be performed on individuals suspected of infection and hence break potential transmission chains. Moreover, in some countries, RT-PCR testing has become a routine procedure of surveillance, testing incoming air passengers in airports \cite{PoolingAirports, TestingAirportsJapan}, schools students \cite{PoolSchool}, or even the entire adult population \cite{MassPooling}.

As such, the need for large-scale testing has
resulted in the application of techniques to increase the throughput
of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
  MatrixPooling, LionDorfman, CherifReview}. These techniques, often
termed \emph{pooling}, or group testing, have originated in the 1940's
in the seminal work of Dorfman and his pooling technique
\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
selects $N$ individuals and performs a single RT-PCR test on their
combined (\emph{pooled}) samples. If the pooled test yields a positive
result, then each individual is retested separately; otherwise,
everyone is declared negative. The throughput efficiency of Dorfman
pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
its error rates thoroughly investigated \cite{Kim, Simplistic1,
  OptimalDorfmanPool}.

@@

\section*{Results}\label{section:results}
\subsection*{Scheme false-negative}
We plot $\err{exact}$ for varying prevalence $q$ and sensitivity $\Se$
values, and make the comparison with $\err{common}$. As recommended by
\cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
different prevalence values. We observe that for a specificity of
$\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
sensitivity and prevalence values \cite{KitComparison,
  InterpretingCOVID19Test, EstimatingRatesLourenco,
  FalsePositiveEstimate}, the increase in the scheme false-negative rate
(compared to $1-\Se$), expressed by $\err{exact}$, is at leaset
$60\%$ (Figure \ref{fig1}).

@@

\section*{Discussion}
In this study, we developed a more realistic and novel probabilistic
model for the outcomes of pooled RT-PCR tests, parametrized for
SARS-CoV-2. Contrary to the common assumption, we assume, based on
data (Table~\ref{table}), that multiple infected individuals increase
the likelihood of a positive pooled test result. A direct consequence
of our model is that low values of infection prevalence increase the
false-negative rates of Dorfman pooling. Importantly, our results
remain qualitatively similar under varying parameter values, in the
observed ranges \cite{KitComparison,EstimatingRatesKucrika,
  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
\ref{fig1}). Therefore, our results give rise to a conflict: low
infection prevalence leads to high efficiency of Dorfman pooling
\cite{DorfmanYuvalDor},but also increases false-negative rates.
